---
title: "A3: SPL Library Checkouts"
Author: Rhea Parida
output: html_document
---

### Introduction

With the advancement of technology, it is said that print is "dying". More people are opting for audio books and ebooks instead of paperbacks and hardcovers. Not only is it the fault of technology, but Ebooks allow people to multitask. They can be listened to in the car, on a walk, or while running errands. One of my main question to answer is if this is true, are people checking out more ebooks and audiobooks than physical books, and to look at this trend over time.   

### Summary Information


The year when the most books were checked out was `r most_books_year` when 1527112 books were checked out. This makes sense because after 2020 there was a large drop in the number of books checked out due to the COVID-19 pandemic. In addition, most people checked out digital material in `r `most_digital_year`, during the pandemic people were hesitant on checking out physical things, let alone go to a public place like the library. Most people checked out `r physica_vs_digital` material between 2017-2023 with 10166737 checkouts. This coincides with the fact that the most popular material type checked out is still `r Most_checkedout_material`, which were checked out 6397967 times. Despite the pandemic causing a drop in physical and book checkouts, these material types and usage class are still very popular among the public. The least popular material type was  `r least_checkedout` with only 10 checkouts. Despite the pandemic making going to the library and checking out physical material hard, it is still amongst the most popular to checkout, which tells us that maybe print is not "dead".   
  



### The Dataset
This data set was collected by the Seattle Public Library. This data set in particular focuses on checkouts from 2017-2023. It lists the checkout type, the material type, the checkout year and month, the number of checkouts, the title of the book, the author, the subject, the publisher, and year of publication. Seattle wants to ensure that that public has access to transparent data, help provide accurate data for future research of economic development, and provide the city with accurate information to help them serve the public more efficiently. Seattle IT collect only the data they need to accomplish their goals, ensure that the public knows that data is being collected, make sure that they follow regulations when collecting sensitive data, and follow a retention plan. The data was collected by the way the books were checked out. We need to make sure that this data cannot be traced back to specific people, and that it is secure. Some possible limitations can include missing information from the data set and there may be inconsistencies in the data.  


### First Trends Over Time Chart


```{r, echo = FALSE, code = readLines("Graph_1.R")}
```

In graph 1 I looked at audiobook, book, and ebook checkouts from 2017 to 2023. Since these are the three ways to get books through the library I wanted to see how popular each material type was throughout the years. From the graph, audiobooks and ebooks have been on a steady rise since 2017. Books on the other hand were checked out more than the other two before 2020, but then dropped after 2020. Books since 2021 however are on the rise, and it looks like they may surpass audiobook and ebook checkouts in the near future. 


### Second Trends Over Time Chart



```{{r, echo = FALSE, code = readLines("graph_2.R")}
```


In my second graph I looked at physical vs. digital checkouts from 2017 to 2023. Physical material includes things like magazines, books, and video disks while digital material includes checkouts of audiobooks, movies, and Ebooks. From this graph up until 2020 people were checking out more physical material than digital material. However, there is a large drop in 2020, likely due to the pandemic, in physical checkouts. Digital material since 2020 has been checked out more than physical material since 2020, but physical material is on a steady rise. 

### Your Choice



```{r, echo = FALSE, code = readLines("Graph_3.R")}
```


For my third graph I pulled all the books by the author Veronica Roth that were checked out over the years and in what format they were checked out in. Out of the 408 observations, none of the books checked out were physical. All of the formats were digital, either through Audiobook or Ebook. Of the two material types, more audiobooks were checked out than ebooks, however there seems to be a spike in Ebook checkouts between 2020 and 2021. This could be because of the Covid-19 pandemic, as people had more time on their hands and were willing to read a book but could not go in person to get a physical copy.  


